Major Studio I \\ Final Project

### 2021-11-05 finalproject/report/finalproject.tex
added report to github

\title{{\LARGE\textbf{7 in 7}} \\ (all the photos are in the images directory, just haven't had time to place them properly in the document)}
\date{\today}
%\author{}

\begin{document}
\maketitle
%\tableofcontents
%----------------------------------------------
\newpage
\subsection*{five (11/03/2021)}
\textbf{name:} trachea support trial no. 1

\textbf{soundtrack:} Ice On The Dune by Empire of the Sun

The objective of this prototype was to create a mechanically robust way to actually make the trachea prototype move and such. First, I did some sketching and head scratching. Then I felt overwhelmed, but I kept thinking. I decided I needed a support on both the curved and flat sides of the trachea tube. One side would just act as support, while the other side allowed the monofillament to be pulled, causing the trachea to contract. After thinking through all kinds of unweidly solutions I had the idea to use pieces of dowel with small holes drilled through both to anchor the trachea and provide the pulling surface.

So I put my shoes on and walked the 80 feet between my apartment and the hardware store -- this apartment is perfect in so many ways -- and got a 3/8" dowel for \$1.40. I measured out a length long enough to suspend the trachea and sit inside a piece of scrap wood and cut two pieces.

Then I marked out where the holes should go; just in line with the vinyl tubing in the trachea. Next, I carefully punched and drilled some small holes through the dowel and added some 3/8" holes -- with a few different spacings -- to my scrap wood to seat the dowels.

The next step was tricky. I needed to create a way to anchor the curved side of the trachea to one of the dowels with holes in it. The catch being no glue was allowed (I alwyas try hard not to use glue when prototyping for maximum material resuse). After thinking for a while, I took a sewing needle and threaded it with monofilament. Then, I carefully pierced and went through the back of the curved side of each vinyl tube leaving about 4" of monofillaent on both sides of the hole.

Finally, I was ready. Starting with the curved, anchor side, I attached one of the dowels to the trachea. Then I seated the anchor dowel in my scrap wood and threaded the other side through the holes. Tying knots in monofillament is always fiddly. The final result looked pretty cool

Then I got to play with it! Wow, it worked better than anticipated!

\begin{figure}[h]
\centering
  \includegraphics[width=0.22\textwidth]{images/46.JPG}
  \includegraphics[width=0.22\textwidth]{images/47.JPG}
  \includegraphics[width=0.22\textwidth]{images/48.JPG}
  \includegraphics[width=0.22\textwidth]{images/49.JPG}
  \includegraphics[width=0.22\textwidth]{images/50.JPG}
  \includegraphics[width=0.22\textwidth]{images/51.JPG}
  \includegraphics[width=0.22\textwidth]{images/52.JPG}
  \caption{sequence of contracting each ring then releasing all of them slowly}
\end{figure}

Takeaways:
\begin{itemize}
\item
it's important to use a sharp needle when going through latex cause it's really stretchy
\item
I need to get this in a position so I can move air through it!
\item
  this works really well for something that's 3" long, but I'll need to use something that's much more ridgid for the 
\end{itemize}
\clearpage
\subsection*{four (11/03/2021)}
\textbf{name:} latex / trachea trial no. 2

@@

After letting this sit for some time, I began to remove the wire pieces and replace them with bits of monofilament like in the previous prototype. I was pleasantly surprised by how well the glue sustained the tension from the vinyl tubing. Nothing began to tear or separate sinificantly, though there was one spot that looked like it didn't get enough glue.

Things I learned:
\begin{itemize}
\item
  this was pretty a easy easy with a 3" section, but I think it will be more difficult with a longer tube
\item
  the latex to latex bond is pretty strong and didn't come apart from the pressure of the viny tubing
\item
  I need to think of a clever way to mount this so that it can be mechanised
\item
  I ended up skimming a cool medical paper about breath noises while making this prototype...need to revisit BUT the gist was that the brachea and trachea are the really important parts of breath noises. The coolest thing I learned -- aside from a bunch of laminar flow stuff in the lungs -- is that our organs and chest stuff act as a low pass filter for noises coming from our lungs and lower brachea / trachea. Most of the higher frequency breath noises that we hear come from the neck where there's less tissue for the sound to travel through. Waves! Definitley need to read more thoracic medicine papers.
\end{itemize}

\clearpage
\section*{nothing (11/02/2021)}

@@

\clearpage
\section*{three (11/01/2021)}
\textbf{name:} latex trial no. 1

@@

\clearpage
\section*{two (10/31/2021)}

@@

To account for these, I think I'm going to try a park. I can't find a louder speaker, but maybe I can make the noise louder digitally. I am also going to create simple interface which will let me adjust which breath sounds are playing instead of having to listen to the youtube video mp3 on repeat.
\clearpage
\subsection*{one (10/30/2021)}

@@

\vspace{1cm}
I think this prototype was pretty successful. It left me with a good number of questions which are as follows.
\begin{itemize}
\item
  how does the length of the output tube effect the sound? need to experiement, but an not ready to cut the tube.
\item
  how does the shape of the cavity effect the sound?
\item
  is this cavity ridiculously large in comparison to normal lungs?
\item
  how do I make the sound less hollow?
\item
  what volume of air is displaced? (if I had a flow meter I could just integrate, but I'll have to be more clever)
\item
  how would a lung structure change the sound?
\item
  how would the sound change if there was some kind of foam or cloth inside the cavity? would that sound less hollow?
\item
  how the heck am I going to automate the diaphragm?
\end{itemize}

\clearpage
\section*{7 in 7 planning}

>>>>

\title{Major Studio I \\ Final Project}
\date{\today}
\author{Lauria Clarke}

\begin{document}
\maketitle
%\tableofcontents
%----------------------------------------------
\section*{the idea}

The Nature of Breath is a machine that breathes where the speed, depth, and "wetness" of the breathing varies in response to COVID 19 data.

Since COVID started we have all experienced a global fascination with breath.

It should be naturalistic in the auditory component and artificial in the visual component.

\section*{mechanical breakdown}

Thinking about the way this device will function mechanically is the hardest part since there are few --- if any -- existing examples. To help think about the systems in a more compartmentalized way, I have broken it wodn into three sub systems: actuation, storage, and modulation. Actuation is the mechanism taht forces air into or outof the system, storage is where air is stored between breaths, and modulation is the subsystem which changes the flow of air and produces the appropriate sound.

In the human body, these three pieces combine to produce breath.

\subsection*{actuation}

In the body, the thing that actuates airflow through the respiratory system is the diaphragm. It contracts causing air ot be firces in or out of the lungs.

Actuation could be achieved in a two ways in this project. First, a changing pressure differential forces air in and out of storage medium. This could be accomplished by changing the pressure in a sealed volume which contains the storage comoponent by either using a diaphragm or directly changing the pressure using a pump. Second, the actuation could be direct -- a bi-directional pump forces air through the airway without any kind of storage component. This solution reduces complexity, but may increase cost.

Either way, the acuation technique needs to be bidirectional -- inhale and exhale.

\subsection*{storage}

Storage isn't necessary if actuation is direct on the airway. For visual purposes however, it would be quite striking to see a shape inflating and deflating with each breath.

\subsection*{modulation}

Modulation is the most complicated and exciting part of this project. It can be furthe briken down into the trachea, larynx, and mouth/sinuses. Each of these three parts of the respiratory system have an impact on how our breating sounds.

In this system, the trachea will be the first point of focus. The trachea consists of C shaped pieces of cartiladge stack on top of each other with a smooth muscle closing the C. When the smooth muscle contracts the trachea grows more narrow allowing us to do things like cough.

% include image of trachea here

This mechanism could pretty simple to replicate.

\section*{initial material gatering and prototypes}

I was quite consumed by this idea after having it, but my ideas for implementation were too flimsy and became quite consuming. I called a number of places that sell balloons and 5 gallon water jug hoping ot find tha materias needed to replicate a gigantic version of that gradeschool experiment. Didn't have much luck. I still couldn't stop thinking about all this, however, so I went to Canal Rubber Co. to look at sheets of latex. This was a GREAT idea. It was soo cool! I gathered more materials than I actually needed and probably spent more than I should've -- \$30.

Then I got home and started playing with the materials and rememebred I'm a bit allergic to latex powder. Ooops!

I build a simple diaphragm test.

\clearpage
\section*{midterm response}

Hi Harpreet, I neglected to read the prompt very thoroughly for this assignment, so what follows is a very unpolished assesment of my current thinking here. More of a journal entry than anything else. My apologies.

After being here (in MFADT) for a couple months I belive I am persuing the following learning goal: to know when to engage in blindly intuitive making and when to engage in research based, highly considered making. So far my life has revolved around intuitive making, an inexplicable itch. Making first and assigning meaning second. I am here, however, to be challenged and that challenge is to become intentional.

On Wednesday, Franco and I went out and had a beer after class. It felt a bit irresponsible from a time management perspective, but it was really helpful to talk about the midterm process totally outside of school while it was so fresh. The bulk of our conversation focused on language. One aspect of Franco's presentation which I really appreciated was the framework he developed to decribe what he was discussing. For me, language is a big focus right now as I think about what it means to be intentional in my work. In order to be intentional we need to posses the language to describe our intent. Finding the appropriate language to use to describe what I'm seeing and trying to achieve feels like the next, most important step in my process.

The easiest place to look for language is in what others have already written. I spent some time looking for this during the year before I came to school, but without knowing the languae to look for and use in my search it was difficult to find a path of work to follow. This was frustrating, because I'm usually pretty good at self directed research and reading. Anyway, I'm definitely in the right place now with the right people to guide me in asking questions and building a foundation of thought for finding answers. Language allows you not only to describe, but question a field.

On the topic of building foundation, I spoke with Tega Brain two weeks ago. I was extremely nervous and practiced a lot. It was a great conversation and she was very generous with her time and knowledge. (As and aside, I am so incredibly excited and grateful be in this place where everyone is willing ot speak with me and discuss the things I'm interested in! I have felt out of place for so much of my academic and professional life it is wonderful to leave taht feeling behind.) I first asked her what frameworks she had encoutnered over her career which have helped lend more specific language and structure to the way the natural and manmade worlds are described and designated. She provided me with many useful references that address that broad ideas I'm trying to learn about.

The conversation naturally flowed into the next question I wanted to ask which was about the similarities between natural optimization processes and manmade or technological optimization procesed. We agreed that the scale of time and the objective of optimization tend to be the distinguishing factors between the two. In the same vein, she described an overarching urge in her work to replicate manmade systems in natural ways, often relying on this idea of optimization. It was helpful to hear that specific sense of purpose stated in such plain language.

Finally I asked her about the feeling of uncertainty which I am trying to investigate which I fell exists in a lot of her work. She asked if I was trying to describe the uncanny valley. I don't think I am, at least not ``the'' uncanny valley. I believe that the unvanny vally is an appropriate technique to describe what I'm talking about, it is accurate in reflecting the feeling of unease, but is specifically describes a human centric condition. What about more broad, natural existence? That is what I'm after. Maybe someone with a better understanding of this question will tell me that the question itself is unnecessary and has already been answered, but that's for me to find out on my own.

Thinking a bit about the review itself, I found the process of preparing the presentation much more helpful than delivering the presentaion. Feedback was certainly useful, but I think I could've structured my presentation to make it more helpful. By framing a few specific questions for the audience to ponder I could've directed the feedback. Definitely need to rememebr that next time. Coming up with questions for others to asses for you impllies that know what help you need with so that you can ask for it. A little circular by definition.

I really liked mark's suggestion of framing the project in diferent ways to understand it better. I'm not sure I entirely know how to go about doing that, but I'm definitly going to try. Richard's coments about following my intuition toward this vein of inquire was helpful in that it made me think more specifically about the sometimes conflicting roles of intuition and intention in this process. The thoughful feedback from class was overwhelming in how geenrous it was (again, feeling lucky to be here)! In many of my classmates' comments I head the same question, which Tanve said really well: what do you want people to take away from your work? Again, a suggestion to focus on my intent.

Overall, the focus on building research and background before engaging in making has made me a lttle bit wary of my own ideas. I think this new degree of consideration is good, but I also recognize that it's important to let an urge to create drive your forward momentum. As I grow as thinkiner and come to understand the space I'm trying to exist in, my process will naturally become more considered. But I've got to start somewhere, and that's right now with the excitement and ideas I have in this moment.

There is a lot of reading ahead of me. Below is an accumulated list from the last month or so.

\section*{accumulated reading list and internet suggestions}

Biomimicry - Janine Benyus (my dad...who has a weird knack for giving me important books at critical moments, like the Toaster Project right when it came out during my senior year of highschool)

seeing is forgetting the name of the thing one sees - Lawrence Welschler (Ed Andrews)

Anthropocene Thesis (not actually the name) - Donna Haraway (Tega Brain)

Against the Anthropocene - T. J. Demos (Tega Brain)

Exhallation - Ted Chang (Tega Brain)

The Sciences of the Artificial - Herbert Simon (Mark ?)

Life's Edge - Carl Zimmer (me)

Making and Being -

\vspace{1cm}
Internet followups from Tega Brain:

\url{http://www.greyroom.org/issues/68/72/the-smartness-mandate-notes-toward-a-critique/}

\url{https://alanwarburton.co.uk/fairytales}

\url{https://jods.mitpress.mit.edu/pub/lewis-arista-pechawis-kite/release/1}

\url{http://www.madlab.cc/}

think this one is the donna haraway\\
\url{https://www.e-flux.com/journal/75/67125/tentacular-thinking-anthropocene-capitalocene-chthulucene/}

\vspace{1cm}
Internet followups from Harpreet:

look at ISEA 2021

\url{https://www.ianingram.org/machines/2009_squirrel.html}

\url{https://www.google.com/search?q=bodies+in+motion&rlz=1C5CHFA_enUS719US719&sxsrf=AOaemvI9jBtAv206EcJ4FUexQurm74VNBw:1633968526778&source=lnms&tbm=isch&sa=X&ved=2ahUKEwiJ7vKM38LzAhXrQ98KHbZjBGoQ_AUoAnoECAIQBA&biw=1440&bih=766&dpr=2}

\url{https://vimeo.com/220895476}

\url{https://news.mit.edu/2018/creating-3-d-printed-motion-sculptures-from-2-d-videos-mit-csail-0919}

\url{https://vvvv.org/blog/bodies-in-motion-humanscale-milan-design-week-2019#:~:text=Bodies%20in%20Motion%20is%20an,scientific%20approach%20to%20furniture%20design}

\url{https://www.youtube.com/watch?v=_v3J6D6R_nI&ab_channel=HumanscaleHQ}

philip beesley

\url{https://www.christianhubert.com/writing/assemblage}

\url{https://link.springer.com/chapter/10.1057/9781137025012_4}

\vspace{1cm}
Internet followups from Richard The:

Design Triennale, Broken Nature

RadioLab -- Breathing podcast

\url{https://www.lozano-hemmer.com/last\_breath.php}

\vspace{1cm}
from Guin: \url{https://www.asianstudies.org/publications/eaa/archives/history-and-sustainability-of-bunraku-the-japanese-puppet-theater/}

from Tanve: The Year the Earth Changed - David Attenborough
%random 
%Building a foundation of language upon which to base your work is critical.
%
%Before coming ot Parsons my practice was veyr much in the vein of making first and assigning meaning second. At present I feel as though I'm being pushed to be more intentional in all areas of my making. 
%
%
%BUIlding a framework or background of language also allows you to call upon references or the prior work in your field more easily. 
%
%I am missing the critical theory that supports the more intuitive exploration I've done so far. I have been looking for it, but without knowing the languae to use in my search it has been difficult to find on my own. A good reminder too that language allows you not only to describe, but question a field. 
%
%I value the feedbak
%
%talk about approaching the work from multiple angles
%I have a hard time letting go of my desire to solve physical challenges through making. This particular desire has very deeply rooted in me for as long as I can remember. Coming up with a physical solution to a complex and often unecessary problem is what drive a lot of my work. 
%
%Again, I think that I need to asses this urge through the lens of intentionality. It can be useful at time, but learning somplementary ways of thinking about the intent or goal of my work is a skil which I really hope to develop in my time here at parsons.
%
%I spoke with tega brain last week and that was a fantastic conversation. She provided me with many, many references that address the questions in her own work and mine. The conversation we had about the specific questions I asked her felt really fruitful as well -- I should've typed this up sooner. We spoke about the paralles between optimization in nature and optimization in engineering and more.
%
%
%chat with Tega Brain
%
%chat with Rob
%
%
%When Mark asked my why ths question of defining what is natural vs. artificial matters I felt like I stumbled a bit. The answers that usually run through my head felt wrong for the moment, but I need to get better at saying what I mean. I have this idealized scenario where a kid grows up in a house where the deck outside is made of fake wood and all the floors in the house are made of fake wood and every single piece of wood around is actually plastic - exactly like Through the Arc of the Rainforest(!). This kid never know what real wood is. Wouldn't that be problematic? To be so out of touch. Maybe I need to re-read through the arc of the rainforest for better insight on this one. Although everyone dies at the end of that book so maybe it's not really so helpful.
%
%
%
%
%suggestion from tanve: what do you want people to take away from your work?

\clearpage
\section*{chat with harpreet on 10/25}

7 in 7 is an attempt to try multiple things in many ways

use different perspective to address the same question

could use

is the conecptual prototype having the intended effect on users

are the users having the intended experience

\clearpage
\section*{{\LARGE7 in 7} (in reverse order, accompanying photos forthcoming)}
also, spellcheck in my text editor isn't working for this document

\newpage
\subsection*{five (11/03/2021)}
\textbf{name:} trachea support trial no. 1

\textbf{soundtrack:} Empire of the Sun

\newpage
\subsection*{four (11/03/2021)}
\textbf{name:} latex / trachea trial no. 2

@@

After letting this sit for some time, I began to remove the wire pieces and replace them with bits of monofilament like in the previous prototype. I was pleasantly surprised by how well the glue sustained the tension from the vinyl tubing. Nothing began to tear or separate.

Things I learned:
\begin{itmemize}
\item
  this was pretty a easy easy with a 3" section, but I think it will be more difficult with a longer tube
\item
  the latex to latex bond is pretty strong and didn't come apart from the pressure of the viny tubing
\item
  I need to think of a clever way to mount this so that it can be mechanised
\item

\end{itemize}

\newpage
\subsection*{nothing (11/02/2021)}

@@

\newpage
\subsection*{three (11/01/2021)}
\textbf{name:} latex trial no. 1

@@

\newpage
\subsection*{two (10/31/2021)}

@@

To account for these, I think I'm going to try a park. I can't find a louder speaker, but maybe I can make the noise louder digitally. I am also going to create simple interface which will let me adjust which breath sounds are playing instead of having to listen to the youtube video mp3 on repeat.
\newpage
\subsection*{one (10/30/2021)}

@@

\vspace{1cm}
I think this prototype was pretty successful. It left me with a good number of questions which are as follows.
\begin{itemize}
\item
  how does the length of the output tube effect the sound? need to experiement, but an not ready to cut the tube.
\item
  how does the shape of the cavity effect the sound?
\item
  is this cavity ridiculously large in comparison to normal lungs?
\item
  how do I make the sound less hollow?
\item
  what volume of air is displaced? (if I had a flow meter I could just integrate, but I'll have to be more clever)
\item
  how would a lung structure change the sound?
\item
  how would the sound change if there was some kind of foam or cloth inside the cavity? would that sound less hollow?
\item
  how the heck am I going to automate the diaphragm?
\end{itemize}
\newpage
\section*{7 in 7 planning}
